---
title: "Vegetation Coverage By Grid"
output: html_document
---

```{r}
library(terra)
library(sf)
```

# Get Coverage by Grid from Orthos

## Merge orthos

For overlapping orthos, merge them together based on their cutlines.

I used the `mosaic` funtion with `blend` from terra here, but I wonder if I should instead use get the max or average values for overlaps, or if I should do more processing in webodm to normalize their reflectance values between orthos. Also experiment with resampling methods - the default bilinear is probably fine, but I could increase it to cubic

```{r}
mergeOrthos <- function(lakeDir){
  imagesList = list.files(lakeDir, pattern=".*\\.tif$")
  if (length(imagesList) > 1){
    # load rasters
    raw <- c()
    for (f in imagesList){
      fpath = file.path(lakeDir, f)
      s <-rast(fpath)
      raw <- c(raw, s)
    }
    out <- do.call(mosaic, (c(raw, fun="blend", resample=TRUE)))
  } else {
    out <- rast(file.path(lakeDir, imagesList[1]))
  }
  return(out)
}
```

## NDVI from Orthos

For each orthophoto, apply preprocessing, then get NDVI from standard formula.

Preprocessing steps:

1.  Mask shadows by filtering by low NIR values, assign these to null?
2.  Remove sun glint with Hedley method

```{r}
getNdviFromOrtho <- function(merged){
  
  
  # clamp values from 0 to 1
  clamped <- clamp(merged, lower=0, upper=1, values=TRUE)
  
  ndvi <- (clamped[[4]] - clamped[[1]]) / (clamped[[4]] + clamped[[1]])
  return(ndvi)
}
```

## Grid over lake

Generate a 10x10m grid covering the image area bounding box. Remove grid cells which don't overlap the lake area. Round the extent coordinates to multiples of 10 (Finnish CRS) so they all align with each other

TODO: restrict this to the image extent? For example for the very large lakes we just did a small area of

```{r}
getLakeGrid <- function(ndvi_bbox){
  # Get corner 
  corner <- st_bbox(ndvi_bbox)[c("xmin", "ymin")]
  # round to ten to align with other grids
  grid_corner <- round(corner/10, 0)*10
  
  lake_grid <- st_make_grid(ndvi_bbox, cellsize=10, offset=grid_corner)
  return (lake_grid)
}
```

## Fractional coverage

Using the following algorithm, get the fractional coverage over the grid.

$$c=\frac{1}{n}\sum_{i=1}^n f_i 
\begin{cases}
    1,  &0.6 \le x_i \\
    \frac{x_i}{(0.6-0.1)}, & 0.1 \le x_i  <  0.6 \\
    0, & x_i < 0.1
\end{cases}$$

```{r}
getFractionalCoverage <- function(ndvi, grid){
  # First apply the classification to the NDVI
  cutoffs <- c(0.1, 0.6)
  classifiedNDVI <- ifel(ndvi < cutoffs[1], 0,
                         ifel(ndvi < cutoffs[2], ndvi / (cutoffs[2] - cutoffs[1]), 1))
  print("classfied NDVI")
  # Now do zonal stats to get mean over the grid
  # TODO: Handle nodata in raster
  zonalOut <- zonal(classifiedNDVI, grid, fun="mean", na.rm=TRUE, as.polygons=TRUE)
  return(zonalOut)
}
```

## Execute functions over all ortho

```{r}
# Define helper function to fall back on the cache.export intermediates if exists and specified
cacheFallback = function(stepFunction, stepArgs, stepOutfile, useCache, exportIntermediates){
  if (useCache & exists(stepOutfile)) {
    # Read from output folder if exists and specified
    stepOutObject <- rast(stepOutfile)
  } else {
    # Else call the function to generate, and export if specified
    stepOutObject <- do.call(stepFunction, list(stepArgs))
    if (exportIntermediates) {
      writeRaster(stepOutObject, stepOutfile, overwrite=TRUE)
      }
  }
  return(stepOutObject)
}

# Define function to export coverage from ortho
exportCoverageFromOrthos <- function(lakedir, useCache=TRUE, exportIntermediates=TRUE, skipCompleted=TRUE){
  outdir = file.path(lakeDir, "outputs")
  dir.create(outdir, showWarnings = FALSE)
  outCoverageGpkg = paste("coverages_",Sys.Date(),".gpkg", sep="")
  outCoverageLayer = basename(lakeDir)
  
  # Skip this layer if specified and the output already exists
  if (skipCompleted & outCoverageLayer %in% st_layers(outCoverageGpkg)$name){
    print(paste("Already completed", outCoverageLayer))
    return()
  } else {
    print(paste("Running", outCoverageLayer))
    # Merge rasters
    merged <- cacheFallback(
      mergeOrthos, 
      c(lakeDir), 
      file.path(outdir, "merged.tif"),
      useCache, exportIntermediates
    )
    print(paste("Merged raster", outCoverageLayer))
    gc()
    
    # Generate NDVI
    # Merge rasters
    ndvi <- cacheFallback(
      getNdviFromOrtho, 
      c(merged), 
      file.path(outdir, "ndvi.tif"),
      useCache, exportIntermediates
    )
    print(paste("Got NDVI", outCoverageLayer))
    gc()
    # Remove merged grid from memory
    rm(merged)
    
    # Generate grid
    ndvi_bbox <- st_as_sf(st_as_sfc(st_bbox(ndvi)))
    grid <- getLakeGrid(ndvi_bbox)
    coverage <- getFractionalCoverage(ndvi, vect(grid))
    print("Got coverage grid")
    
    # Write coverage grid to gpkg
    writeVector(
      coverage, 
      filename=outCoverageGpkg, 
      layer=outCoverageLayer,
      overwrite=TRUE
    )
    gc()
    return()
  }
}
```

```{r}
# Remember to mount "new volume" first
orthos_path = "/media/northross/New Volume/Orthos"
lakeDirs.list = list.dirs(orthos_path, recursive=FALSE)[2:26]

# Set to TRUE to used cached images if they exist
useCache = TRUE

# Set to true to skip a file if its final output already exists
skipCompleted = TRUE
# And whether to export intermediate files
exportIntermediates = TRUE

for (lakeDir in lakeDirs.list) {
  exportCoverageFromOrthos(lakeDir, useCache, exportIntermediates, skipCompleted)
}
```

# Classify Grid coverage

Spatial join the grid to the polygons generated from plant points and lines. If only one species is present in the grid cell, set that species to 1 and all others to zero.

```{r}
lake_outlines = st_read('/home/northross/Documents/Thesis_Data/InitialAnalysis/SurveyLakes.gpkg', layer="SurveyLakes")
```
