---
title: "Vegetation Coverage By Grid"
output: html_document
---

```{r}
library(terra)
library(sf)
```

# Get Coverage by Grid from Orthos

## Merge orthos

For overlapping orthos, merge them together based on their cutlines.

I used the `mosaic` funtion with `blend` from terra here, but I wonder if I should instead use get the max or average values for overlaps, or if I should do more processing in webodm to normalize their reflectance values between orthos. Also experiment with resampling methods - the default bilinear is probably fine, but I could increase it to cubic

```{r}
mergeOrthos <- function(dir){
  imagesList = list.files(dirs.list[1], pattern="*.tif")
  if (length(imagesList) > 1){
    # load rasters
    raw <- c()
    for (f in imagesList){
      fpath = file.path(dir, f)
      s <-rast(fpath)
      raw <- c(raw, s)
    }
    out <- do.call(mosaic, (c(raw, fun="blend", resample=TRUE)))
  } else {
    out <- rast(imagesList[1])
  }
  return(out)
}
```

## NDVI from Orthos

For each orthophoto, correct reflectances \<0 to zero for all bands, then get NDVI from standard formula.

```{r}
getNdviFromOrtho <- function(merged){
  # clamp values from 0 to 1
  corrected <- clamp(merged, lower=0, upper=1, values=TRUE)
  ndvi <- (corrected[[4]] - corrected[[1]]) / (corrected[[4]] + corrected[[1]])
  return(ndvi)
}
```

## Grid over lake

Generate a 10x10m grid covering the image area bounding box. Remove grid cells which don't overlap the lake area. Round the extent coordinates to multiples of 10 (Finnish CRS) so they all align with each other

```{r}
getLakeGrid <- function(lake_outlines, dir){
  # Get all lakes with group id matching the current dir
  dir_num <- as.integer(gsub("\\D", "", dir))
  sel_lakes <- lake_outlines[lake_outlines$groupID == dir_num,]
  
  # Get corner 
  corner <- st_bbox(sel_lakes)[c("xmin", "ymin")]
  # round to ten to align with other grids
  grid_corner <- round(corner/10, 0)*10
  
  lake_grid <- st_make_grid(sel_lakes, cellsize=10, offset=grid_corner)
  return (lake_grid)
}
```

## Fractional coverage

Using the following algorithm, get the fractional coverage over the grid.

$$c=\frac{1}{n}\sum_{i=1}^n f_i 
\begin{cases}
    1,  &0.6 \le x_i \\
    \frac{x_i}{(0.6-0.1)}, & 0.1 \le x_i  <  0.6 \\
    0, & x_i < 0.1
\end{cases}$$

```{r}
getFractionalCoverage <- function(ndvi, grid){
  
}
```

## Execute functions over all ortho

```{r}
dirs.list = list.dirs('Orthos')[2:26]
lake_outlines = st_read('/home/northross/Documents/Thesis_Data/InitialAnalysis/SurveyLakes.gpkg', layer="SurveyLakes")

for (dir in dirs.list) {
  merged <- mergeOrthos(dir)
  ndvi <- getNdviFromOrtho(merged)
  # export intermediates?
  grid <- getLakeGrid(lake_outlines, dir)
  
  coverage <- getFractionalCoverage(ndvi, grid)
  # export coverage
}
```

# Classify Grid coverage

Spatial join the grid to the polygons generated from plant points and lines. If only one species is present in the grid cell, set that species to 1 and all others to zero.

Next, find the nearest point or line feature and set the

```{r}

```
